---
title: "Elastic Net (EN) Analysis Report"
output:
  html_document:
    toc: true
    theme: default
    highlight: textmate
---

# Setup 
```{r showCode, echo=FALSE}
showCode <- TRUE
```

## Knitr Options 
This code provides options for the report. 
```{r knitrSetup, purl=FALSE, echo=showCode}
library(knitr)

# Knitr default options
opts_chunk$set(dpi=300, tidy=TRUE, fig.align='center', echo=showCode)
```

## Load Elastic Net Functions and Data
This code loads all functions necessary function to run elastic net and data.
```{r setup}
# Set random seed
set.seed(1)

# Load functions
source(file.path(.lmp, "ElasticNetExps", "src", "ElasticNet.R"))
source(file.path(.lmp, "ElasticNetExps", "src", "plotElasticNet.R"))  
source(file.path(.lmp, "Utils", "src", "GenUtils.R"))

# Load data
load(file.path(.lmp, "RData", "lmpdb.Rdata"))

# Set variables that will be used for elastic net
molDB <- elNetMolDataNCI60
drugActData <- drugDB$act
```

# Set Elastic Net Parameters
These are the parameters used for the elastic net run, including: 

* Drug to be analyzed
* Cell lines to be used in analysis
* CellMiner datasets to be used
* Restrictions on genes 
* Thresholds for feature filtering
* Thresholds for outputs

```{r setEnParam}
# Human Readable Name
analysisName <- ""

# Drug NSC
drug <- "609699"

# Excluded features and cell lines
excludedCellLines <- NULL # DOES NOT WORK, DO NOT TRY
excludeDatasets <- c("mir", "pro", "mda")

#excludeFeatures <- c("expSLFN11")
excludeFeatures <- NULL

## Restrict gene set 
load(file.path(.lmp, "gene_set_pathway_analysis", "data", "ddr_genelist.Rdata"))
#restrictedGeneSet <- rownames(ddr.list)
restrictedGeneSet <- NULL

# Show debugging data?
verbose <- TRUE

# Thresholds 
corThreshold <- 0.3
minNumResponsiveLines <- 3
responseThreshold <- 0.5

# Results parameters
numFeaInResults <- 5
maxPlotPredictors <- 50

outputFilePath <- "enResults.csv"
```

# Elastic Net
## Set Up Result Storage
(Documentation TBA)
```{r setupResults, message=FALSE}
numDrugs <- length(drug)
resultTypes <- c("Name", "Wt", "CumCor")

# Make empty data.frame of the right size
batchElNetResults <- as.data.frame(matrix(NA, nrow=numDrugs, 
                                          ncol=numFeaInResults*length(resultTypes)), stringsAsFactors=FALSE)

tmpNames <- NULL
colCnt <- 1

for(i in 1:numFeaInResults) {
  for(j in 1:length(resultTypes)) {
    tmpName <- paste("Feature", i, resultTypes[j], sep="")
    tmpNames <- c(tmpNames, tmpName)
    
    # Set data.frame column data type
    if(j %% length(resultTypes) == 1) {
      batchElNetResults[,colCnt] <- vector(mode="character", length=numDrugs)
    } else {
      batchElNetResults[,colCnt] <- vector(mode="numeric", length=numDrugs)      
    }  
  }  
  
  colCnt <- colCnt + 1
}

names(batchElNetResults) <- tmpNames
rownames(batchElNetResults) <- drug
```

## Run Elastic Net
### Setup Feature Set

This code filters the features used in the elastic net analysis. 

```{r setupFeatureSet}
tic()
# Validate the existence of drugs in drug activity data
stopifnot(all(is.element(drug, rownames(drugActData))))

drugActProfile <- as.numeric(drugActData[drug, ])
names(drugActProfile) <- colnames(drugActData)

# Dataset Indicies
datasetIdx <- !(names(molDB) %in% excludeDatasets)

# SKIP DRUG ACTIVITY PROFILES WITH AN INSUFFICIENT NUMBER OF RESPONSIVE LINES.
numResponsive <- sum(drugActProfile > responseThreshold, na.rm=TRUE)
if (numResponsive < minNumResponsiveLines){
  next
}

# EXCLUDE CELL LINES WITH MISSING ACTIVITY DATA AND RECORD VALID CELL LINES 
# TO ACCESS CORRESPONDING MOLECULAR DATA.
drugActProfile <- drugActProfile[!is.na(drugActProfile) & !(names(drugActProfile) %in% excludedCellLines)]
validCellLines <- names(drugActProfile)

# Filter datasets
featureSet <- selectCorrelatedRows2(Y=drugDB$act[drug, ], 
                                    XList=elNetMolDataNCI60[datasetIdx], 
                                    corThreshold=0.3)

# Filter features 
if(!is.null(restrictedGeneSet)) {
  featureSet <- restrictFeatureMat(geneSet=restrictedGeneSet, featureMat=featureSet)  
}

featureSet <- featureSet[setdiff(rownames(featureSet), excludeFeatures), ]
```

### Run Elastic Net 
```{r runEn}
# RUN ELASTIC NET
# ADJUSTED CALL TO RECORD CUMULATIVE CORRELATIONS FOR EXPANDED SET OF PREDICTORS.  
elNetResults <- elasticNet(featureMat=featureSet, responseVec=drugActProfile, 
                           alphaVals=seq(0.6, 0.8, length=10), 
                           cumCorNumPredictors=maxPlotPredictors, verbose=FALSE)
toc()
```

# Results
## EN Run Summary

* Feature Sets Used: `r paste(names(molDB)[datasetIdx], collapse=", ")`
* Features Used: `r dim(featureSet)[1]`
* Samples Used: `r dim(featureSet)[2]`
* Includes expSLFN11: `r if(is.null(excludeFeatures)) {TRUE} else {FALSE}`
* All Genes: `r if(is.null(restrictedGeneSet)) {TRUE} else {FALSE}`
* Alpha: `r elNetResults$alpha` 
* Lambda: `r elNetResults$lambda`


### Mean-Cross Validation Error Plot

This plot shows potential models at different values of lambda: 

* X-axis: Values of lambda tested
* Y-axis: Prediction error of model
* Values across top: Size (number of features) for each model at different values of lambda
* Left dashed line: Indicates the lambda value with the minimum prediction error
* Right dashed line: Indicates the lambda value with a prediction error one standard deviation from the minimum

```{r enRunInfo}
plot(elNetResults$cvOutput)
```

## EN Results Plot

This plot shows 3 pieces of information for the resulting model:

* A heatmap of values for each feature
* The weights of each features
* The drug activity for each sample

```{r plotResults, fig.width=8, fig.height=8}
# Parameters used in EN plot
numPlotPredictors <- min(length(elNetResults$predictorWts), maxPlotPredictors)
featureAnnot <- genFeatureAnnotations(molDB, elNetResults, drugActProfile)

plotElasticNet(drugName=drug,
               weights=elNetResults$predictorWts[1:numPlotPredictors], 
               drugAct=drugActData, 
               molDb=molDB, 
               numCellLines=length(validCellLines), 
               numFeatures=numPlotPredictors,
               featureAnnotations=featureAnnot,
               showCellLineLabels=TRUE, 
               pdfFilename=NULL, 
							 thresholdValues=TRUE,
							 verbose=TRUE)
```

# Save Data

This code saves the results of the elastic net analysis.

```{r writeRawData}
# Save elastic net results object
save(elNetResults, file=paste0(drug, ".Rdata"))

# Generate summary table
for(i in 1:numFeaInResults) {
  for(j in 1:length(resultTypes)) {
    if(resultTypes[j] == "Name") {
      batchElNetResults[drug, paste("Feature", i, resultTypes[j], sep="")] <- names(elNetResults$predictorWts[i])        
    }
    
    if(resultTypes[j] == "Wt") {
      batchElNetResults[drug, paste("Feature", i, resultTypes[j], sep="")] <- elNetResults$predictorWts[i]       
    }
    
    if(resultTypes[j] == "CumCor") {
      batchElNetResults[drug, paste("Feature", i, resultTypes[j], sep="")] <- elNetResults$predictedResponseCor[i]       
    }
  }
}

write.csv(batchElNetResults, file=outputFilePath)

fileName <- paste(drug, ".elnetdata.csv", sep="")
load(paste0(drug, ".Rdata"))

rowNames <- c(drug, names(elNetResults$predictorWts))
elnetData <- matrix(0, nrow=length(rowNames), ncol=ncol(drugActData))
rownames(elnetData) <- rowNames
colnames(elnetData) <- colnames(drugActData)

elnetData[drug, ] <- as.numeric(drugActData[drug, ])

for (predName in names(elNetResults$predictorWts)){
  elnetData[predName, ] <- molDB[[getMolDataType(predName)]][predName, ]
}

elnetData <- as.data.frame(elnetData)

write.csv(x=elnetData, file=fileName, quote=FALSE)
```

# Session Information

## Bitbucket Mercurial Repository Information

Below is the revision identifier for the code used.

```{r hgInfo}
system(paste0("hg id ", .lmp), intern=TRUE)
```

## R Session Information

Below are the versions of R and packages used in this analysis.

```{r sessionInfo}
sessionInfo()
```
