---
title: "Elastic Net (EN) Analysis Report"
output:
  html_document:
    toc: true
    theme: default
    highlight: textmate
---

# Setup 
This code provides options for the report. 
```{r knitrSetup, purl=FALSE, echo=TRUE}
library(knitr)

# Knitr default options
opts_chunk$set(dpi=150, tidy=TRUE, fig.align='center', echo=TRUE)
```

## Load Elastic Net Functions and Data
This code loads all functions and data needed to run elastic net.
```{r setup, message=FALSE}
# Set random seed
set.seed(1)

library(rcellminer)
library(rcellminerData)
library(rcellminerElasticNet)
library(impute)
library(glmnet)

molDB <- getMolDataMatrices()
drugActData <- exprs(getAct(rcellminerData::drugData))
```

# Set Elastic Net Parameters
These are the parameters used for the elastic net run, including: 

* Drug to be analyzed
* Cell lines to be used in analysis
* CellMiner datasets to be used
* Restrictions on genes 
* Thresholds for feature filtering
* Thresholds for outputs
```{r setEnParam}
# Drug NSC
drug <- "<%= nsc %>"

# Included datasets
includeDatasets <- <%= includeDatasets %>
includeFeaturesPath <- '<%= includeFeaturesPath %>'

# Excluded features and cell lines
excludeCellLines <- <%= excludeCellLines %> 
#excludeFeatures <- c("expSLFN11")
excludeFeatures <- NULL

## Restrict gene set 
# load(file.path(.lmp, "gene_set_pathway_analysis", "data", "ddr_genelist.Rdata"))
# restrictedGeneSet <- rownames(ddr.list)
restrictedGeneSet <- NULL

# Use intersection of features across data types in molDB.
useCommonFeaturesInMolDB <- <%= useCommonFeaturesInMolDB %>

# Show debugging data?
verbose <- <%= verbose %>

# THRESHOLDS 
# Features must have absolute correlation greater than corThreshold to be
# included in elastic net.
corThreshold <- <%= corThreshold %>
# Elastic net results will be augmented to include information about input features 
# that are correlated with selected predictors above predictorCorFeatureThreshold.
predCorFeatureThreshold <- <%= predCorFeatureThreshold %>
minNumResponsiveLines <- <%= minNumResponsiveLines %>
responseThreshold <- <%= responseThreshold %>

minNumMutations <- <%= minNumMutations %>

# Results parameters
maxPlotPredictors <- <%= maxPlotPredictors %>

# Determines if intercept term is to be used in predicting held out response
# values during cross validation.  This should be set to FALSE for leave-out-out
# cross validation; TRUE setting can be evaluated for 5 or 10 fold CV.
useModelYIntercept <- <%= useModelYIntercept %>

filePrefix <- "<%= filePrefix %>"
outputDir <- "<%= outputDir %>"
outputFilePath <- "enResults.csv"
```

# Elastic Net
## Setup Feature Set

This code prepares the data used in the elastic net analysis. 

```{r setupFeatureSet, message=FALSE}
tic()

#----[prepare drug data]-----------------------------------------------------------------
# Make sure data is available for drug.
stopifnot(all(is.element(drug, rownames(drugActData))))

# Exclude selected cell lines.
drugActData <- drugActData[, -which(colnames(drugActData) %in% excludeCellLines)]

# Exclude cell lines w/o activity data for drug of interest
drugActProfile <- setNames(as.numeric(drugActData[drug, ]), colnames(drugActData))
drugActProfile <- drugActProfile[!is.na(drugActProfile)]
drugActData <- drugActData[, names(drugActProfile)]

# NOTE DRUG ACTIVITY PROFILES WITH AN INSUFFICIENT NUMBER OF RESPONSIVE LINES.
numResponsive <- sum(drugActProfile > responseThreshold, na.rm=TRUE)
if (numResponsive < minNumResponsiveLines){
  filePrefix <- paste0("invalid_", filePrefix)
}

#----[prepare molecular data]------------------------------------------------------------
# Restrict molecular data to drug data cell line set prepared above.
molDB <- lapply(molDB, function(x) {
  x[, names(drugActProfile)]
})

# Restrict molecular data to specified subtypes.
datasetIdx <- (names(molDB) %in% includeDatasets)
molDB <- molDB[datasetIdx]

# Impute missing gene expression data. 
if ("exp" %in% names(molDB)){
  # NOTE: invisible() is used to suppress function print statements
  molDB[["exp"]] <- invisible(impute.knn(molDB[["exp"]])$data)
}

# Transform to z-score data.
if ("swa" %in% names(molDB)){
  molDB[["swa"]] <- t(scale(t(molDB[["swa"]])))
}

# Filter mutation profiles with fewer than minNumMutations.
if ("mut" %in% names(molDB)){
  countMutations <- function(mutPattern){
    return(sum(mutPattern > 0))
  }

  mutCounts <- apply(molDB$mut, MARGIN = 1, FUN = countMutations)
  molDB$mut <- molDB$mut[(mutCounts >= minNumMutations), ]
  
  # Transform to z-score data.
  molDB[["mut"]] <- t(scale(t(molDB[["mut"]])))
}

#----[prepare elastic net feature data]--------------------------------------------------

if (includeFeaturesPath != ""){
  includeFeatures <- read.table(includeFeaturesPath, header = TRUE, stringsAsFactors = FALSE)
  includeFeatures <- includeFeatures[,1]
  for (molDBType in names(molDB)) {
    filteredFeatures <- intersect(rownames(molDB[[molDBType]]), paste0(molDBType, includeFeatures))
    molDB[[molDBType]] <- molDB[[molDBType]][filteredFeatures, ]
  }
}

if (useCommonFeaturesInMolDB){
  molDBFeatures <- lapply(molDB, function(x) {unname(removeMolDataType(rownames(x)))})
  commonFeatures <- Reduce(intersect, molDBFeatures)
  for (molDBType in names(molDB)){
    molDB[[molDBType]] <- molDB[[molDBType]][paste0(molDBType, commonFeatures), ]
  }
}

# Filter datasets
featureSet <- selectCorrelatedRowsFromMatrices(Y=drugActProfile, 
                                               XList=molDB, 
                                               corThreshold=corThreshold)

# Filter features 
if(!is.null(restrictedGeneSet)) {
  featureSet <- restrictFeatureMat(geneSet=restrictedGeneSet, featureMat=featureSet, 
                                   prefixSet=includeDatasets)  
}

featureSet <- featureSet[setdiff(rownames(featureSet), excludeFeatures), ]

# Included feature sets
names(molDB)
```

## Run Elastic Net 
```{r runEn, message=FALSE}
# RUN ELASTIC NET
# ADJUSTED CALL TO RECORD CUMULATIVE CORRELATIONS FOR EXPANDED SET OF PREDICTORS. 
tryCatch({
  elNetResults <- elasticNet(featureMat=featureSet, responseVec=drugActProfile, id=drug,
                           alphaVals=seq(<%= alphaValMin %>, <%= alphaValMax %>, length = 10), 
                           useLambdaMin=TRUE, cumCorNumPredictors=maxPlotPredictors, 
                           verbose=verbose, useOneStdErrRule = TRUE,
                           standardize=TRUE, standardizeY=TRUE, fitIntercept=FALSE)
}, error=function(e) {
  elNetResults <- list()
  filePrefix <- paste0("invalid_", filePrefix)
  save(elNetResults, file=file.path(outputDir, paste0(filePrefix, ".Rdata"))) 
  stop(e)
})

# NOTE: Add features that correlated to resulting predictors
#elNetResults <- addPredictorCorrelatedFeatures(elNetResults, molDB[datasetIdx], predCorFeatureThreshold)

toc()
```

# Results
## EN Run Summary

* NSC: <%= nsc %>
* Feature Sets Used: `r paste(names(molDB), collapse=", ")`
* Features Used: `r dim(featureSet)[1]`
* Samples Used: `r dim(featureSet)[2]`
* Alpha: `r elNetResults$alpha` 
* Lambda: `r elNetResults$lambda`

### EN Feature Summary 
The table below provides feature counts for each of the feature sets used. 
```{r enFeaSummary}
table(substr(rownames(featureSet), 1, 3))
```

### Mean-Cross Validation Error Plot

This plot shows potential models at different values of lambda: 

* X-axis: Values of lambda tested
* Y-axis: Prediction error of model
* Values across top: Size (number of features) for each model at different values of lambda
* Left dashed line: Indicates the lambda value with the minimum prediction error
* Right dashed line: Indicates the lambda value with a prediction error one standard deviation from the minimum

```{r enRunInfo, eval=FALSE}
# No longer applicable with repeats of cross-validation.
# plot(elNetResults$cvOutput)
```

## EN Results Plot

This plot shows 3 pieces of information for the resulting model:

* A heatmap of values for each feature
* The weights of each features
* The drug activity for each sample
```{r plotResults, fig.width=8, fig.height=8}

numPlotPredictors <- min(length(elNetResults$predictorWts), maxPlotPredictors)
featureAnnot <- genFeatureAnnotations(molDB, elNetResults, as.numeric(drugActData[drug, ]))
featureAnnot <- featureAnnot[1:numPlotPredictors]

plotElasticNet(drugName=drug,
               weights=elNetResults$predictorWts[1:numPlotPredictors], 
               drugAct=drugActData, 
               molDb=molDB, 
               numCellLines=length(drugActProfile), 
               numFeatures=numPlotPredictors,
               featureAnnotations=featureAnnot,
               showCellLineLabels=TRUE, 
               pdfFilename=NULL, 
							 thresholdValues=TRUE,
               topMarWt=2.4, 
               bottomMarWt=1.5,
               leftMarHeatmap=6,
							 verbose=FALSE)
```

An explanation of the predictor annotations to the left of the above heatmap:

- **cr** indicates the correlation of the predictor with the drug response (shown below heatmap).
- **cm** indicates the correlation of the weighted sum of predictors, up to and including
the indicated one, with the drug response.
- **fq** indicates the selection frequency for the predictor over data subset repeat runs.


Cross validation results for elastic net model (feature set) selection.
```{r cvResults, fig.width=8, fig.height=8}
iMinEstErr <- which.min(elNetResults$cvMeanSqErrVals)
if (length(iMinEstErr) == 1){
  maxErr <- elNetResults$cvMeanSqErrVals[iMinEstErr] + elNetResults$cvSdMeanSqErrVals[iMinEstErr]
} else{
  maxErr <- NA
}
k <- length(elNetResults$predictorWts)

x <- seq_along(elNetResults$cvMeanSqErrVals)
y <- elNetResults$cvMeanSqErrVals
se <- elNetResults$cvSdMeanSqErrVals

plot(x, y, xlab="Number of Predictors", ylab="Cross-Validation Error",
     main="Cross-Validation Errors for Nested Predictor Sets",
     ylim=c(0, max(1, max(y+se, na.rm = TRUE))))

if (!is.na(maxErr)){
  abline(h = maxErr, col="red")
}
abline(v=k, col="green", lty=3, lwd=2.5)
segments(x,y-se,x,y+se)
epsilon <- 0.10
segments(x-epsilon, y-se, x+epsilon, y-se)
segments(x-epsilon, y+se, x+epsilon, y+se)

x <- seq_along(elNetResults$cvPredRsqVals)
y <- elNetResults$cvPredRVals
plot(x, y, xlab="Number of Predictors", ylab="Pearson's Correlation",
     main="CV-Predicted vs. Actual Response Correlation",
     ylim=c(0, 1))
abline(v=k, col="green", lty=3, lwd=2.5)
```

## Run ridge regression using (pre-model selection) elastic net predictors. 
```{r runRidgeWithEnPreds, message=FALSE}
tryCatch({
  if (length(elNetResults) > 0){
    enPredsPreModSelection <- names(elNetResults$predictorWts_preModelSelection)
    if (length(enPredsPreModSelection) > 1){
      set.seed(1)
      enPredMat <- t(featureSet[enPredsPreModSelection, , drop = FALSE])
      ridgeCvOut <- glmnet::cv.glmnet(x = enPredMat, y = drugActProfile, alpha = 0)
      ridgeOut <- glmnet::glmnet(x = enPredMat, y = drugActProfile, alpha = 0)
      elNetResults$predictorWts_enSelectedRidge <- predict(ridgeOut, 
        type = "coefficients", s = ridgeCvOut$lambda.min)[-1, ]
    }
  }
})
```

Cross-validation of overall elastic net.
```{r fullEnCvResults, fig.width=8, fig.height=8}

# Leave-one-out cross-validation.
set.seed(2)
cvFoldIds <- getCvFoldIds(nObs = length(drugActProfile), nFolds = length(drugActProfile))

elNetResults$fullEnCvResults <- cvElasticNet( 
  cvFoldIds=cvFoldIds, keepEnResults=FALSE,
  featureMat=featureSet, responseVec=drugActProfile, id=drug,
  alphaVals=seq(<%= alphaValMin %>, <%= alphaValMax %>, length = 10), useLambdaMin=TRUE,
  cumCorNumPredictors=maxPlotPredictors, verbose=verbose, 
  useModelYIntercept = useModelYIntercept,
  useOneStdErrRule = TRUE,
  standardize=TRUE, standardizeY=TRUE, fitIntercept=FALSE)

# Correlation between full EN cross-validation predicted values and actual values.
elNetResults$fullEnCvResults$cvPredR

cvEnPredTab <- elNetResults$fullEnCvResults$cvFoldEnPredictorTab[
  c("PREDICTOR", "FREQUENCY", "SIGN")]
cvEnPredTab <- cvEnPredTab[order(cvEnPredTab$FREQUENCY, decreasing = TRUE), ]

kable(cvEnPredTab, caption="Predictors Selected During CV Elastic Net Runs", 
      row.names=FALSE, digits=2)
```

# Save Data
This code saves the results of the elastic net analysis.

```{r saveData}
save(elNetResults, file=file.path(outputDir, paste0(filePrefix, ".Rdata"))) 
```

# Session Information
```{r versionInfo}
cat("CV Run Time: ", sum(elNetResults$fullEnCvResults$enRunTimes), "\n")

cat("Timestamp: ", format(Sys.time(), "%m%d%yT%H%M%S"), "\n")
sessionInfo()  
```
