---
title: "Report"
output:
  html_document:
    toc: true
    theme: default
    highlight: textmate
---

# Setup 
This code provides options for the report. 
```{r knitrSetup, purl=FALSE, echo=TRUE}
library(knitr)

# Knitr default options
opts_chunk$set(dpi=150, tidy=TRUE, fig.align='center', echo=TRUE)
```

## Load Elastic Net Functions and Data
This code loads all functions and data needed to run elastic net.
```{r setup, message=FALSE}
# Set random seed
set.seed(1)

library(rcellminer)
library(rcellminerData)
library(rcellminerElasticNet)
library(impute)

molDB <- getMolDataMatrices()
drugActData <- exprs(getAct(rcellminerData::drugData))
```

# Set Elastic Net Parameters
These are the parameters used for the elastic net run, including: 

* Drug to be analyzed
* Cell lines to be used in analysis
* CellMiner datasets to be used
* Restrictions on genes 
* Thresholds for feature filtering
* Thresholds for outputs
```{r setEnParam}
# Drug NSC
drug <- "<%= nsc %>"

# Included datasets
includeDatasets <- <%= includeDatasets %>

# Excluded features and cell lines
excludeCellLines <- <%= excludeCellLines %> 
#excludeFeatures <- c("expSLFN11")
excludeFeatures <- NULL

## Restrict gene set 
# load(file.path(.lmp, "gene_set_pathway_analysis", "data", "ddr_genelist.Rdata"))
# restrictedGeneSet <- rownames(ddr.list)
restrictedGeneSet <- NULL

# Show debugging data?
verbose <- TRUE

# THRESHOLDS 
# Features must have absolute correlation greater than corThreshold to be
# included in elastic net.
corThreshold <- <%= corThreshold %>
# Elastic net results will be augmented to include information about input features 
# that are correlated with selected predictors above predictorCorFeatureThreshold.
predCorFeatureThreshold <- <%= predCorFeatureThreshold %>
minNumResponsiveLines <- <%= minNumResponsiveLines %>
responseThreshold <- <%= responseThreshold %>

minNumMutations <- <%= minNumMutations %>

# Results parameters
maxPlotPredictors <- <%= maxPlotPredictors %>

filePrefix <- "<%= filePrefix %>"
outputDir <- "<%= outputDir %>"
outputFilePath <- "enResults.csv"
```

# Results
## EN Run Summary

* NSC: <%= nsc %>

### EN Feature Summary 

```{r}
y <- drugActData["94600",]
x <- molDB[["exp"]]["expSLFN11",]
plot(x, y, main="<%= nsc %>")
```

# Session Information
```{r versionInfo}
cat("Timestamp: ", format(Sys.time(), "%m%d%yT%H%M%S"), "\n")
sessionInfo()  
```
